Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-26 NEUROC/sec4-experiments.tex
更新正文。

\figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} further compares the best accuracy that every GNN achieved on different datasets.
%
The best accuracy of the four GNNs was very close.
%
%
It was also close to the accuracy reported in \cite{}.
%
The results indicated that there was no clear winner.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
GaAN achieved the highest accuracy in three out of five datasets.
%
GCN achieved the highest or second-highest accuracy in three out of five datasets, though its model was simplest.
%
Simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

@@

The complexity analysis in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} was valid for both training and inference.
%
Fixing other hyper-parameters, each hyper-parameter itself affected the training time and the memory usage of a GNN Layer \emph{in a linear way}.
%
Algorithm engineers could adjust hyper-parameters according to the time complexity to avoid explosive growth in the training/inference time and memory usage.
%
The accuracy of GNNs was much more sensitive to the dimension of hidden vectors/heads than the other hyper-parameters.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
There was no clear winner for all datasets.

\subsection{Training/Inference Time Breakdown}
\label{sec:training_time_breakdown}

To find out which stage/step dominated the training time, we decomposed the training time and analyzed performance bottlenecks level by level.

\subsubsection{Layer Level}

@@

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gaan.pdf}}
    \caption{Training time breakdown on the layer level. The training time of each layer included the time spent on the forward, backward, and evaluation phases. Each layer was further decomposed into the vertex and the edge calculation stages.}
    \label{fig:exp_vertex_edge_cal_proportion}
\end{figure}

Each GNN layer was further divided into the vertex and the edge calculation stages.
%
In \figurename~\ref{fig:exp_vertex_edge_cal_proportion}, GCN spent most of the training time on the edge calculation stage on most datasets.
%
A special case was the \texttt{cph} dataset.
%
The dimension of the input feature vectors was very high in \texttt{cph}, making the vertex calculation stage of the GCN Layer 0 spend considerable time.
%
GGNN also spent the majority of its training time on the edge calculation stage, but the high time complexity of its vertex updating function $\gamma^l$ made the proportion of the vertex calculation in the total training time much higher than the other GNNs.
%
For GAT and GaAN, due to their high edge calculation complexity, the edge calculation stage was the dominant stage.

@@

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gaan.pdf}}
    \caption{Effects of the average degree on the time proportion of the edge/vertex calculation. Graphs were generated with the R-MAT generator by fixing the number of vertices as 50,000. }
    \label{fig:exp_avg_degree_on_vertex_edge_cal_time}
\end{figure}

@@

The results indicate that \emph{the performance bottlenecks of the edge calculation stage depend on the complexity of the messaging function $\phi^l$}.
%
When the time complexity of $\phi^l$ was high, the messaging step was the performance bottleneck.
%
Optimizing the implementation of $\phi$ could significantly reduce training time.
%
Otherwise, the collection and the aggregation steps were performance bottlenecks.
%
Improving the efficiency of the two steps could benefit all GNNs.

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_edge_calc_decomposition_gaan.pdf}}
    \caption{Training time breakdown of the edge calculation stage.}
    \label{fig:exp_edge_calc_decomposition}
\end{figure}

\subsubsection{Operator Level}

The functions $\phi$, $\Sigma$ and $\gamma$ in the edge and vertex calculation stages were made up of a series of basic operators implemented on the GPU side, such as the matrix multiplication \texttt{mm}, the elementwise multiplication \texttt{mul} and the index-based selection \texttt{index\_select}.
%
\figurename~\ref{fig:exp_top_basic_ops} shows the top-5 time-consuming basic operators in each GNN, averaged over all datasets.

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_top_basic_ops_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_top_basic_ops_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_top_basic_ops_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_top_basic_ops_gaan.pdf}}
    \caption{Top 5 time-consuming basic operators of typical GNNs. The time proportion of each basic operator was averaged over all datasets with the error bar indicating the maximum and the minimum.}
    \label{fig:exp_top_basic_ops}
\end{figure}

@@

%
We found that the performance characteristics of GNN inference were highly similar to training on the \emph{layer} level and the \emph{step} level.
%
Taking GCN as an example, \figurename~\ref{fig:inference_analysis_on_layer_edge_level_of_gcn} shows the results of GCN inference on the \emph{layer} level and the \emph{step} level.
%
By cross-comparing \figurename~\ref{fig:exp_vertex_edge_cal_proportion}(a), 
\figurename~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time}(a), and \figurename~\ref{fig:exp_edge_calc_decomposition}(a),
we could conclude that the time breakdowns were very similar.
%
For the other GNNs, we observed similar phenomena.
%

% GCN Inference结果对比图
\begin{figure}[H]
    \centering
    \subfloat[Layer Level]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[Effects of the average degree]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[Edge Level]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_inference_full_edge_calc_decomposition_gcn.pdf}}
    \caption{Time breakdowns on the layer level and edge level, and effects of average degree on the edge/vertex calculation of GCN inference.}
    \label{fig:inference_analysis_on_layer_edge_level_of_gcn}
\end{figure}

% 绝对时间对比图
\begin{figure}[H]
    \centering
    \subfloat[\texttt{amc}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-computers.pdf}}
    \subfloat[\texttt{amp}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-photo.pdf}}
    \subfloat[\texttt{fli}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_flickr.pdf}}
    \caption{Wall-clock inference time on different datasets.}
    \label{fig:compare_wall_clock_time_of_training_and_inference}
\end{figure}

@@

The top time-consuming basic operators of training and inference also showed a certain degree of difference.
%
Some of the top time-consuming operators during training were replaced by new operators in inference.
%
Figure~\ref{fig:exp_inference_top_basic_ops} in the response compares the top 5 time-consuming basic operators in inference.
%
For GCN, the \texttt{index} operator used in the prediction layer became the new top 5 time-consuming operators, replacing the \texttt{gather} operator used in the backward phase in training.
%
%    The basic operators related to the edge calculation (\texttt{scatter\_add}, \texttt{index\_select}, and \texttt{mul}) still consumed the majority of the inference time.
%
For GGNN, \texttt{index} operator in the prediction layer also became one of the top 5 time-consuming basic operators.
%
%    Due to the high time complexity of the vertex updating function $\gamma$ in GGNN, the basic operators related to the vertex calculation stage (\texttt{mm} and \texttt{thnn\_fused\_gru\_cell}) still consumed near half of the inference time.
%
For GAT, the \texttt{input\_put\_impl} operator (used in the backward phase) in Figure~\ref{fig:exp_top_basic_ops}(c) was replaced by the \texttt{scatter\_add} operator used in the forward edge calculation in Figure~\ref{fig:exp_inference_top_basic_ops}(c).
%
%    The \texttt{scatter\_add} operator was related to the edge calculation step.
%
For GaAN, the \texttt{mm} operator was replaced by the \texttt{cat} operator used in the vertex updating function.

@@

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gaan.pdf}}
    \caption{Top 5 time-consuming basic operators of typical GNNs. The time proportion of each basic operator was averaged over all datasets with the error bar indicating the maximum and the minimum.}
    \label{fig:exp_inference_top_basic_ops}
\end{figure}

\paragraph{Summary}
% TODO:
The GNN training/inference was suitable for GPUs.
%
\emph{The edge calculation stage was the main performance bottleneck in most cases}, except for training GNNs with high vertex calculation complexity on low-average-degree graphs.
%
The performance bottleneck in the edge calculation stage depended on the time complexity of the messaging function $\phi$.
%
If the time complexity of $\phi$ was {high}, {$\phi$} dominated the training time of the edge calculation stage.
%
Optimizations should focus on improving its efficiency.
%
Otherwise, the {collection step} and the {aggregation step} dominated the training time.
%
The collection step suffered from lots of data movement.
%
The aggregation step suffered from data synchronization and non-regular data access.

@@

@@

The average degree of the sampled subgraph was \emph{much lower} than the average degree of the original graph, especially when the relative batch size was low.
%
Taking the neighbor sampler with the relative batch size of 6\% as an example, the average degree of the \texttt{amp} dataset was 31.1, but the average degree of the sampled subgraph was only 5.8.
%
For the cluster sampler, the average degree was 3.0.
%
\figurename~\ref{fig:exp_sampling_minibatch_degrees_distribution} compares the degree distribution of the sampled subgraphs with the original graph.
%
The slopes of the curves were similar, indicating that the sampled subgraphs still followed the power-law degree distribution.
%
However, the numbers of high-degree vertices were much less than the original graph, lowering the average degrees.
%
According to the experimental results in Section~\ref{sec:training_time_breakdown}, if the average degree became lower, the proportion of the training time spent on the vertex calculation stage would become higher, especially for GGNN.

>>>>

\figurename~\ref{fig:exp_hyperparameter_on_accuracy_alg_contrast} further compares the best accuracy that every GNN achieved on different datasets.
%
The best accuracy of the four GNNs was very close.
%
%
It was also close to the accuracy reported in \cite{shchur2018_pitfall_of_gnn, zeng2020_graphsaint}.
%
The results indicated that there was no clear winner.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
GaAN achieved the highest accuracy in three out of five datasets.
%
GCN achieved the highest or second-highest accuracy in three out of five datasets, though its model was simplest.
%
Simple GNN models (such as GCN) could still achieve good accuracy with proper hyper-parameter settings.

@@

The complexity analysis in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} was valid for both training and inference.
%
Fixing other hyper-parameters, each hyper-parameter itself affected the training/inference time and the memory usage of a GNN layer \emph{in a linear way}.
%
Algorithm engineers could adjust hyper-parameters according to the complexity analysis to avoid explosive growth in training/inference time and memory usage.
%
The accuracy of GNNs was much more sensitive to the dimensions of hidden vectors/heads than the other hyper-parameters.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
Simple GNN models (such as GCN) could achieve good accuracy with proper hyper-parameter settings.

\subsection{Time Breakdown Analysis}
\label{sec:time_breakdown_analysis}

To find out which stage/step dominated the training/inference time, we decomposed the training/inference time and analyzed performance bottlenecks level by level.
%
We first analyzed the training time in Section~\ref{sec:time_breakdown_analysis_layer_level} to Section~\ref{sec:time_breakdown_analysis_operator_level}.
%
The similarities and differences between training and inference were analyzed in Section~\ref{sec:time_breakdown_analysis_inference}.

\subsubsection{Layer Level}
\label{sec:time_breakdown_analysis_layer_level}

@@

\begin{figure}[H]
    \centering
    \subfloat[GCN\label{fig:exp_vertex_edge_cal_proportion_gcn}]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4.6cm]{figs/experiments/exp_layer_time_proportion_gaan.pdf}}
    \caption{Training time breakdown on the layer level. The training time of each layer included the time spent on the forward, backward, and evaluation phases. Each layer was further decomposed into the vertex and the edge calculation stages.}
    \label{fig:exp_vertex_edge_cal_proportion}
\end{figure}

Each GNN layer was further divided into the vertex and edge calculation stages.
%
In \figurename~\ref{fig:exp_vertex_edge_cal_proportion}, GCN spent most of the training time on the edge calculation stage on most datasets.
%
A special case was the \texttt{cph} dataset.
%
The dimension of the input feature vectors was very high in \texttt{cph}, making the vertex calculation stage of the GCN Layer 0 spend considerable time.
%
GGNN also spent the majority of its training time on the edge calculation stage, but the high time complexity of its vertex updating function $\gamma^l$ made the proportion of the vertex calculation in the total training time much higher than the other GNNs.
%
For GAT and GaAN, due to their high edge calculation complexity, the edge calculation stage was the dominant stage.

@@

\begin{figure}[H]
    \centering
    \subfloat[GCN\label{fig:exp_avg_degree_on_vertex_edge_cal_time_gcn}]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gaan.pdf}}
    \caption{Effects of average degrees on the vertex/edge calculation time. Graphs were generated with the R-MAT generator by fixing the number of vertices as 50,000. }
    \label{fig:exp_avg_degree_on_vertex_edge_cal_time}
\end{figure}

@@

The results indicate that \emph{the performance bottlenecks of the edge calculation stage depend on the complexity of the messaging function $\phi$}.
%
When the time complexity of $\phi$ was high, the messaging step was the performance bottleneck.
%
Optimizing the implementation of $\phi$ could significantly reduce training time.
%
Otherwise, the collection and the aggregation steps were performance bottlenecks.
%
Improving the efficiency of the two steps could benefit all GNNs.

\begin{figure}[H]
    \centering
    \subfloat[GCN\label{fig:exp_edge_calc_decomposition_gcn}]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_edge_calc_decomposition_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_edge_calc_decomposition_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_edge_calc_decomposition_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_edge_calc_decomposition_gaan.pdf}}
    \caption{Training time breakdown of the edge calculation stage.}
    \label{fig:exp_edge_calc_decomposition}
\end{figure}

\subsubsection{Operator Level}
\label{sec:time_breakdown_analysis_operator_level}

The functions $\phi$, $\Sigma$ and $\gamma$ in the edge and vertex calculation stages were made up of a series of basic operators implemented on the GPU side, such as the matrix multiplication \texttt{mm}, the elementwise multiplication \texttt{mul} and the index-based selection \texttt{index\_select}.
%
\figurename~\ref{fig:exp_top_basic_ops} shows the top-5 time-consuming basic operators of training each GNN, averaged over all datasets.

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_top_basic_ops_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_top_basic_ops_ggnn.pdf}}\\
    \subfloat[GAT\label{fig:exp_top_basic_ops_gat}]{\includegraphics[height=4cm]{figs/experiments/exp_top_basic_ops_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_top_basic_ops_gaan.pdf}}
    \caption{Top 5 time-consuming basic operators in training. The time proportion of each basic operator was averaged over all datasets with the error bar indicating the maximum and the minimum.}
    \label{fig:exp_top_basic_ops}
\end{figure}

@@

\subsubsection{Comparing Inference and Training}
\label{sec:time_breakdown_analysis_inference}

%
We found that the performance characteristics of GNN inference were highly similar to training on the \emph{layer} level and the \emph{step} level.
%
Taking GCN as an example, \figurename~\ref{fig:inference_analysis_of_gcn} shows the time breakdowns of GCN inference on the \emph{layer} level and \emph{step} level of the edge calculation.
%
By cross-comparing \figurename~\ref{fig:exp_vertex_edge_cal_proportion_gcn} with \figurename~\ref{fig:inference_analysis_of_gcn_layer_level}, 
\figurename~\ref{fig:exp_avg_degree_on_vertex_edge_cal_time_gcn} with \figurename~\ref{fig:inference_analysis_of_gcn_avg_degree}, and \figurename~\ref{fig:exp_edge_calc_decomposition_gcn} with \figurename~\ref{fig:inference_analysis_of_gcn_edge_level}, the time breakdowns of training and inference were very similar.
%
For the other GNNs, we observed similar phenomena.

\begin{figure}[H]
    \centering
    \subfloat[Layer level\label{fig:inference_analysis_of_gcn_layer_level}]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_inference_full_layer_time_proportion_gcn.pdf}}
    \subfloat[Effects of average degrees\label{fig:inference_analysis_of_gcn_avg_degree}]{\includegraphics[width=0.4\columnwidth]{figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}\\
    \subfloat[Step level of the edge calculation stage\label{fig:inference_analysis_of_gcn_edge_level}]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_inference_full_edge_calc_decomposition_gcn.pdf}}
    \caption{Time breakdowns of GCN inference.}
    \label{fig:inference_analysis_of_gcn}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \subfloat[\texttt{amc}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-computers.pdf}}
    \subfloat[\texttt{amp}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_amazon-photo.pdf}}
    \subfloat[\texttt{fli}]{\includegraphics[width=0.3\columnwidth]{figs/experiments/exp_time_comparison_between_training_inference_flickr.pdf}}
    \caption{Wall-clock inference time on different datasets.}
    \label{fig:compare_wall_clock_time_of_training_and_inference}
\end{figure}

The top time-consuming basic operators of training and inference also showed a certain degree of difference.
%
Some of the top time-consuming operators during training were replaced by new operators in inference.
%
Figure~\ref{fig:exp_inference_top_basic_ops} shows the top 5 time-consuming basic operators in inference.
%
For GCN, the \texttt{index} operator used in the prediction layer became the new top 5 time-consuming operators, replacing the \texttt{gather} operator used in the backward phase in training.
%
%    The basic operators related to the edge calculation (\texttt{scatter\_add}, \texttt{index\_select}, and \texttt{mul}) still consumed the majority of the inference time.
%
For GGNN, \texttt{index} operator in the prediction layer also became a time-consuming basic operators.
%
%    Due to the high time complexity of the vertex updating function $\gamma$ in GGNN, the basic operators related to the vertex calculation stage (\texttt{mm} and \texttt{thnn\_fused\_gru\_cell}) still consumed near half of the inference time.
%
For GAT, the \texttt{input\_put\_impl} operator (used in the backward phase) in Figure~\ref{fig:exp_top_basic_ops_gat} was replaced by the \texttt{scatter\_add} operator used in the forward edge calculation in Figure~\ref{fig:exp_inference_top_basic_ops_gat}.
%
%    The \texttt{scatter\_add} operator was related to the edge calculation step.
%
For GaAN, the \texttt{mm} operator was replaced by the \texttt{cat} operator used in the vertex updating function.

@@

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_ggnn.pdf}}\\
    \subfloat[GAT\label{fig:exp_inference_top_basic_ops_gat}]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_inference_full_top_basic_ops_gaan.pdf}}
    \caption{Top 5 time-consuming basic operators in inference. The time proportion of each basic operator was averaged over all datasets with the error bar indicating the maximum and the minimum.}
    \label{fig:exp_inference_top_basic_ops}
\end{figure}

\paragraph{Summary of Time Breakdown Analysis}

The GNN training/inference was suitable for GPUs.
%
{The edge calculation stage was the main performance bottleneck in most cases}, except for processing GNNs with high vertex calculation complexity on low-average-degree graphs.
%
The performance bottlenecks in the edge calculation stage depended on the time complexity of the messaging function $\phi$.
%
If the time complexity of $\phi$ was {high}, {$\phi$} dominated the training/inference time of the edge calculation stage.
%
Optimizations should focus on improving its efficiency.
%
Otherwise, the {collection step} and the {aggregation step} dominated the training/inference time.
%
The collection step suffered from lots of data movement.
%
The aggregation step suffered from data synchronization and non-regular data access.

@@

\paragraph{Training}

@@

The average degree of the sampled subgraph was \emph{much lower} than the average degree of the original graph, especially when the relative batch size was low.
%
Taking the neighbor sampler with the relative batch size of 6\% as an example, the average degree of the \texttt{amp} dataset was 31.1, but the average degree of the sampled subgraph was only 5.8.
%
For the cluster sampler, the average degree was 3.0.
%
\figurename~\ref{fig:exp_sampling_minibatch_degrees_distribution} compares the degree distribution of the sampled subgraphs with the original graph.
%
The slopes of the curves were similar, indicating that the sampled subgraphs still followed the power-law degree distribution.
%
However, the numbers of high-degree vertices were much less than the original graph, lowering the average degrees.
%
According to the experimental results in Section~\ref{sec:time_breakdown_analysis}, if the average degree became lower, the proportion of the training time spent on the vertex calculation stage would become higher, especially for GGNN.
